---
output: html_document
---

```{r, echo=FALSE}
library(dplyr, quietly=TRUE)

all_samples = read.table("input/all_samples.txt", sep="\t", header=TRUE, encoding="UTF-8", stringsAsFactors=FALSE)
all_samples$Randomisation[all_samples$Randomisation==""] = "Exclus"

all_samples$Progression = ifelse(is.na(all_samples$Prog_V06) & is.na(all_samples$Prog_V12), NA,
                              ifelse(!is.na(all_samples$Prog_V06) & all_samples$Prog_V06=="VRAI", "V06",
                                ifelse(all_samples$Prog_V12=="VRAI", "V12", "Aucune")))


rna_samples = read.table("input/rna_samples.txt", sep="\t", header=TRUE, encoding="UTF-8", stringsAsFactors=FALSE)
```

# Summarize by gleason score 

Here are some tables summarizing available samples at V00 and V06 by their Gleason score, 
and whether or not we have RNA samples available for one or both time points.

## For all samples:
### V00

```{r, echo=FALSE}
samples_by_gleason = all_samples %>% 
    filter(P189_ToutesBiopsies_Visite.PCa14=="V00") %>% 
    group_by(StudyArmCode) %>% 
    summarize(G6=sum(BioGleason==6 | is.na(BioGleason), na.rm=TRUE),
              G7=sum(BioGleason==7, na.rm=TRUE),
              G8=sum(BioGleason==8, na.rm=TRUE))
knitr::kable(samples_by_gleason)
```

### V06
```{r, echo=FALSE}
samples_by_gleason = all_samples %>% 
    filter(P189_ToutesBiopsies_Visite.PCa14=="V06") %>% 
    group_by(StudyArmCode) %>% 
    summarize(G6=sum(BioGleason==6 | is.na(BioGleason), na.rm=TRUE),
              G7=sum(BioGleason==7, na.rm=TRUE),
              G8=sum(BioGleason==8, na.rm=TRUE))
knitr::kable(samples_by_gleason)
```


## For samples with RNA available.
### V00

```{r, echo=FALSE}
samples_by_gleason = all_samples %>% 
    filter(P189_ToutesBiopsies_Visite.PCa14=="V00" & (ARN_V00_NONA=="VRAI" | ARN_V06_NONA=="VRAI")) %>% 
    group_by(StudyArmCode) %>% 
    summarize(G6=sum(BioGleason==6 | is.na(BioGleason), na.rm=TRUE),
              G7=sum(BioGleason==7, na.rm=TRUE),
              G8=sum(BioGleason==8, na.rm=TRUE))
knitr::kable(samples_by_gleason)
```

### V06

```{r, echo=FALSE}
samples_by_gleason = all_samples %>% 
    filter(P189_ToutesBiopsies_Visite.PCa14=="V06" & (ARN_V00_NONA=="VRAI" | ARN_V06_NONA=="VRAI")) %>% 
    group_by(StudyArmCode) %>% 
    summarize(G6=sum(BioGleason==6 | is.na(BioGleason), na.rm=TRUE),
              G7=sum(BioGleason==7, na.rm=TRUE),
              G8=sum(BioGleason==8, na.rm=TRUE))
knitr::kable(samples_by_gleason)
```

## For samples with *matched* RNA available.
### V00
```{r, echo=FALSE}
samples_by_gleason = all_samples %>% 
    filter(P189_ToutesBiopsies_Visite.PCa14=="V00" & ARN_V00_NONA=="VRAI" & ARN_V06_NONA=="VRAI") %>% 
    group_by(StudyArmCode) %>% 
    summarize(G6=sum(BioGleason==6 | is.na(BioGleason), na.rm=TRUE),
              G7=sum(BioGleason==7, na.rm=TRUE),
              G8=sum(BioGleason==8, na.rm=TRUE))
knitr::kable(samples_by_gleason)
```

### V06
```{r, echo=FALSE}
samples_by_gleason = all_samples %>% 
    filter(P189_ToutesBiopsies_Visite.PCa14=="V06" & ARN_V00_NONA=="VRAI" & ARN_V06_NONA=="VRAI") %>% 
    group_by(StudyArmCode) %>% 
    summarize(G6=sum(BioGleason==6 | is.na(BioGleason), na.rm=TRUE),
              G7=sum(BioGleason==7, na.rm=TRUE),
              G8=sum(BioGleason==8, na.rm=TRUE))
knitr::kable(samples_by_gleason)
```

# Summarize patients by progression

Here are tables summarizing how many patients progressed from a low grade to a high-grade cancer.
The timing of the progression (whether it was detected at V06 or V12) is also tracked.
To get the total number of patients that progressed at V12, add the Progression 6 mois 
and Progression 12 mois columns.

## For all samples
```{r, echo=FALSE}
first_patient_rows = all_samples %>% filter(!duplicated(IDPCa14_1))

samples_by_gleason = first_patient_rows %>% 
    group_by(StudyArmCode) %>% 
    summarize("Non applicable"=sum(is.na(Progression)),
              "Aucune progression"=sum(Progression=="Aucune", na.rm=TRUE),
              "Progression 6 mois"=sum(Progression=="V06", na.rm=TRUE),
              "Progression 12 mois"=sum(Progression=="V12", na.rm=TRUE))
knitr::kable(samples_by_gleason)

prog_df = samples_by_gleason %>% 
            mutate(ProgAll=`Progression 6 mois`+ `Progression 12 mois`) %>%
            mutate(Total=`Aucune progression` + ProgAll) %>%
            mutate(PercentProg=ProgAll/Total)
diet_row = prog_df$StudyArmCode=="Diète"
diet_prog = prog_df$ProgAll[diet_row]
diet_no_prog = prog_df[["Aucune progression"]][diet_row]
diet_total = diet_prog + diet_no_prog
diet_percent = sprintf("%2.1f", diet_prog / diet_total * 100)
ari_row = prog_df$StudyArmCode=="5-ARI"
ari_prog = prog_df$ProgAll[ari_row]
ari_no_prog = prog_df[["Aucune progression"]][ari_row]
ari_total = ari_prog + ari_no_prog
ari_percent = sprintf("%2.1f", ari_prog / ari_total * 100)
diet_under = phyper(diet_prog, diet_prog + ari_prog, diet_no_prog + ari_no_prog, diet_total)
```

There are `r diet_prog`/`r diet_total` (`r diet_percent`%) progressions at month 12 for the diet arm, and
`r ari_prog`/`r ari_total` (`r ari_percent`%) progressions at month 12 in the anti-androgen arm. The p-value
of underrepresentation of progressions in the diet arm as assessed by Fischer test is 
`r sprintf("%1.2f", diet_under)`.

## For samples with RNA available.
```{r, echo=FALSE}
samples_by_gleason = first_patient_rows %>% 
    filter(ARN_V00_NONA=="VRAI" | ARN_V06_NONA=="VRAI") %>% 
    group_by(StudyArmCode) %>% 
    summarize("Non applicable"=sum(is.na(Progression)),
              "Aucune progression"=sum(Progression=="Aucune", na.rm=TRUE),
              "Progression 6 mois"=sum(Progression=="V06", na.rm=TRUE),
              "Progression 12 mois"=sum(Progression=="V12", na.rm=TRUE))
knitr::kable(samples_by_gleason)
```


## For samples with matched RNA available.
```{r, echo=FALSE}
samples_by_gleason = first_patient_rows %>% 
    filter(ARN_V00_NONA=="VRAI" & ARN_V06_NONA=="VRAI") %>% 
    group_by(StudyArmCode) %>% 
    summarize("Non applicable"=sum(is.na(Progression)),
              "Aucune progression"=sum(Progression=="Aucune", na.rm=TRUE),
              "Progression 6 mois"=sum(Progression=="V06", na.rm=TRUE),
              "Progression 12 mois"=sum(Progression=="V12", na.rm=TRUE))
knitr::kable(samples_by_gleason)
```

