---
output: html_document
---

```{r, echo=FALSE}
library(dplyr, quietly=TRUE)

all_samples = read.table("input/all_samples.txt", sep="\t", header=TRUE, encoding="UTF-8", stringsAsFactors=FALSE)
all_samples$Randomisation[all_samples$Randomisation==""] = "Exclus"

all_samples$Progression = ifelse(is.na(all_samples$Prog_V06) & is.na(all_samples$Prog_V12), NA,
                              ifelse(!is.na(all_samples$Prog_V06) & all_samples$Prog_V06=="VRAI", "V06",
                                ifelse(all_samples$Prog_V12=="VRAI", "V12", "Aucune")))


rna_samples = read.table("input/rna_samples.txt", sep="\t", header=TRUE, encoding="UTF-8", stringsAsFactors=FALSE)
```

# Summarize by gleason score 

## For all samples:
### V00

```{r, echo=FALSE}
samples_by_gleason = all_samples %>% 
    filter(P189_ToutesBiopsies_Visite.PCa14=="V00") %>% 
    group_by(Randomisation) %>% 
    summarize(G6=sum(BioGleason==6 | is.na(BioGleason), na.rm=TRUE),
              G7=sum(BioGleason==7, na.rm=TRUE),
              G8=sum(BioGleason==8, na.rm=TRUE))
knitr::kable(samples_by_gleason)
```

### V06
```{r, echo=FALSE}
samples_by_gleason = all_samples %>% 
    filter(P189_ToutesBiopsies_Visite.PCa14=="V06") %>% 
    group_by(Randomisation) %>% 
    summarize(G6=sum(BioGleason==6 | is.na(BioGleason), na.rm=TRUE),
              G7=sum(BioGleason==7, na.rm=TRUE),
              G8=sum(BioGleason==8, na.rm=TRUE))
knitr::kable(samples_by_gleason)
```


## For samples with RNA available.
### V00

```{r, echo=FALSE}
samples_by_gleason = all_samples %>% 
    filter(P189_ToutesBiopsies_Visite.PCa14=="V00" & ARN=="VRAI") %>% 
    group_by(Randomisation) %>% 
    summarize(G6=sum(BioGleason==6 | is.na(BioGleason), na.rm=TRUE),
              G7=sum(BioGleason==7, na.rm=TRUE),
              G8=sum(BioGleason==8, na.rm=TRUE))
knitr::kable(samples_by_gleason)
```

### V06

```{r, echo=FALSE}
samples_by_gleason = all_samples %>% 
    filter(P189_ToutesBiopsies_Visite.PCa14=="V06" & ARN=="VRAI") %>% 
    group_by(Randomisation) %>% 
    summarize(G6=sum(BioGleason==6 | is.na(BioGleason), na.rm=TRUE),
              G7=sum(BioGleason==7, na.rm=TRUE),
              G8=sum(BioGleason==8, na.rm=TRUE))
knitr::kable(samples_by_gleason)
```

## For samples with *matched* RNA available.
### V00
```{r, echo=FALSE}
samples_by_gleason = all_samples %>% 
    filter(P189_ToutesBiopsies_Visite.PCa14=="V00" & ARN_V00_NONA=="VRAI" & ARN_V06_NONA=="VRAI") %>% 
    group_by(Randomisation) %>% 
    summarize(G6=sum(BioGleason==6 | is.na(BioGleason), na.rm=TRUE),
              G7=sum(BioGleason==7, na.rm=TRUE),
              G8=sum(BioGleason==8, na.rm=TRUE))
knitr::kable(samples_by_gleason)
```

### V06
```{r, echo=FALSE}
samples_by_gleason = all_samples %>% 
    filter(P189_ToutesBiopsies_Visite.PCa14=="V06" & ARN_V00_NONA=="VRAI" & ARN_V06_NONA=="VRAI") %>% 
    group_by(Randomisation) %>% 
    summarize(G6=sum(BioGleason==6 | is.na(BioGleason), na.rm=TRUE),
              G7=sum(BioGleason==7, na.rm=TRUE),
              G8=sum(BioGleason==8, na.rm=TRUE))
knitr::kable(samples_by_gleason)
```

# Progression
## For all samples
```{r, echo=FALSE}
samples_by_gleason = all_samples %>% 
    filter(P189_ToutesBiopsies_Visite.PCa14=="V00") %>% 
    group_by(Randomisation) %>% 
    summarize("Non applicable"=sum(is.na(Progression)),
              "Aucune progression"=sum(Progression=="Aucune", na.rm=TRUE),
              "Progression 6 mois"=sum(Progression=="V06", na.rm=TRUE),
              "Progression 12 mois"=sum(Progression=="V12", na.rm=TRUE))
knitr::kable(samples_by_gleason)

prog_df = samples_by_gleason %>% 
            mutate(ProgAll=`Progression 6 mois`+ `Progression 12 mois`) %>%
            mutate(Total=`Aucune progression` + ProgAll) %>%
            mutate(PercentProg=ProgAll/Total)
diet_row = grepl("Avodart 6 mois", prog_df$Randomisation)
diet_prog = prog_df$ProgAll[diet_row]
diet_no_prog = prog_df[["Aucune progression"]][diet_row]
diet_total = diet_prog + diet_no_prog
diet_percent = sprintf("%2.1f", diet_prog / diet_total * 100)
ari_row = grepl("Avodart 12 mois", prog_df$Randomisation)
ari_prog = prog_df$ProgAll[ari_row]
ari_no_prog = prog_df[["Aucune progression"]][ari_row]
ari_total = ari_prog + ari_no_prog
ari_percent = sprintf("%2.1f", ari_prog / ari_total * 100)
diet_under = phyper(diet_prog, diet_prog + ari_prog, diet_no_prog + ari_no_prog, diet_total)
```

There are `r diet_prog`/`r diet_total` (`r diet_percent`%) progressions for the diet arm, and
`r ari_prog`/`r ari_total` (`r ari_percent`%) progressions in the anti-androgen arm. The p-value
of underrepresentation of progressions in the diet arm is `r sprintf("%1.2f", diet_under)`.

## For samples with RNA available.
```{r, echo=FALSE}
samples_by_gleason = all_samples %>% 
    filter(P189_ToutesBiopsies_Visite.PCa14=="V00" & ARN=="VRAI") %>% 
    group_by(Randomisation) %>% 
    summarize("Non applicable"=sum(is.na(Progression)),
              "Aucune progression"=sum(Progression=="Aucune", na.rm=TRUE),
              "Progression 6 mois"=sum(Progression=="V06", na.rm=TRUE),
              "Progression 12 mois"=sum(Progression=="V12", na.rm=TRUE))
knitr::kable(samples_by_gleason)
```


## For samples with matched RNA available.
```{r, echo=FALSE}
samples_by_gleason = all_samples %>% 
    filter(P189_ToutesBiopsies_Visite.PCa14=="V00" & ARN_V00_NONA=="VRAI" & ARN_V06_NONA=="VRAI") %>% 
    group_by(Randomisation) %>% 
    summarize("Non applicable"=sum(is.na(Progression)),
              "Aucune progression"=sum(Progression=="Aucune", na.rm=TRUE),
              "Progression 6 mois"=sum(Progression=="V06", na.rm=TRUE),
              "Progression 12 mois"=sum(Progression=="V12", na.rm=TRUE))
knitr::kable(samples_by_gleason)
```

